Constant-round Blind Classical Verification of Quantum Sampling

### 2020-11-14 introduction.tex
Sync

\section{Introduction}
% \XW{
% \begin{itemize}
%     \item add a lot of references.
%     \item comparison with the most relevant results:
%       \begin{itemize}
%           \item Sampling, the only paper; how about classical sampling?
%           \item the following for BQP  
%           \item blind and verifiable ~\cite{GV19}; we  constant round; technique-wise very different.
%           \item there is a table in~\cite{Grilo19}. Safe to say ~\cite{GV19} only existing blind protocol in the computational setting?
%           \item all previous either quantum clients, or at least 2 provers.
%           \item the following for blindness
%           \item Mahadev in her thesis~\cite{mahadev_2018} discussed a bit about the relation between verifiability and blindness. She hoped to get verifiability out of blindness by designing some non-malleable QFHE but failed.   
%           \item what's the high-level message we can say here?  Use QFHE in a different way? It is correct that not much work in the classical setting either.  Maybe existing classical work employs the principle but with different implementation.
%           \item old approach, first get blindness  (measurement-based, self-testing), and then try to make it verifiable; our approach, first have a verifiable protocol, and upgrade by a QFHE.
%           \item directly QFHE (blindness) won't give verifiability. some thoughts from Mahadev.
%           \item
%       \end{itemize}
% \end{itemize}
% }

Can quantum computation, with potential computational advantages that are intractable for classical computers,
be efficiently verified by classical clients?
This seeming paradox has been one of the central problems in quantum complexity theory and delegation of quantum computation~\cite{web:Aaronson}.
From a philosophical point of view, this question is also known to have a fascinating connection to the \emph{falsifiability} of quantum mechanics in the potential high complexity regime~\cite{survey:AV12}.

A complexity theoretic formulation of this problem by Gottesman in 2004~\cite{web:Aaronson} asks the possibility for an efficient classical verifier/client (a $\BPP$ machine) to verify the output of an
efficient quantum prover (a $\BQP$ machine).
In the absence of techniques for directly tackling this question, earlier feasibility results on this problem have been focusing on two weaker formulations.
The first type of feasibility results (e.g.,~\cite{BFK09,arXiv:ABOEM17,FK17,mf16}) considers the case where the $\BPP$ verifier is equipped with limited quantum power.
The second type of feasibility results (e.g,~\cite{Nat:RUV13, CGJV19, Gheorghiu_2015, HPF15})
considers a $\BPP$ verifier interacting with at least two entangled, non-communicating quantum provers.
In a recent breakthrough, Mahadev~\cite{FOCS:Mahadev18a} proposed the first protocol of classical verification of quantum computation (CVQC) whose soundness is based on a widely recognized computational assumption that the learning with error (LWE) problem~\cite{JACM:Regev09} is hard for $\BQP$ machines.
The technique invented therein has inspired many  subsequent developments of CVQC protocols with improved parameters and functionality (e.g.,~\cite{FOCS:GheVid19,arXiv:AlaChiHun19,arXiv:ChiaChungYam19}).
On the other side, there are known complexity-theoretic limitations on the feasibility of blind CVQC in the information-theoretical setting (e.g.~\cite{aaronson_et_al:LIPIcs:2019:10582}).
%\XW{here some no-go for information theoretical result.}
%We refer curious readers to a slightly outdated  survey~\cite{survey:GKK19} for details.

@@

It is worthwhile noting that there is a simple reduction of the delegation of \emph{classical} sampling problems to decision ones: for example, the verifier can fix a seed for pseudo-randomness, send it to the prover, and then the sampling outcome can be computed  bit-by-bit under this pseudo-random seed in a deterministic way. As such, classical literature of delegation of computation primarily focuses on delegation of decision problems. %delegation of classical sampling problem is not explicitly studied in literature.

Unfortunately, it is unclear about how to make this de-randomization trick work for the delegation of $\SampBQP$, 
as randomness in quantum computation inherently comes from quantum mechanics rather than classical probability theory. 
Thus, it seems that the delegation of $\SampBQP$ needs to take a different technical route.
%delegation of $\SampBQP$ may require a totally different construction from delegation of $\BQP$. 
%
One natural starting point is Mahadev's CVQC protocol~\cite{FOCS:Mahadev18a} for $\BQP$. 
Indeed, some nice features of Mahadev's protocol (e.g., allowing X-Z measurements on any qubit) suggest the feasibility of generating  multi-bit measurement outcomes with the protocol. 
However, as highlighted below in the technical introduction section, there are also important drawbacks of Mahadev's protocol in its current form, which makes it hard to allow a CVQC protocol for $\SampBQP$ with desired performance. As we shall see, such difficulties do not hold for the decision problem, which demonstrates an important difference between CVQC protocols for the decision and the sampling problems.

% \XW{Adjust the flow; highlight the key questions}
% \Ethan{This paragraph is new and has to be checked}
% Even with the CVQC protocol for $\BQP$ from \cite{FOCS:Mahadev18a}, however, 
%it is still unclear how to construct a CVQC protocol for $\SampBQP$.
% In the context of delegating \emph{classical} sampling problems,
% there is a simple reduction to that of decision problems.
% Namely, the verifier fixes a seed for pseudorandomness and send it to the prover,
% then the sampling outcome can be computed deterministically bit-by-bit using this pseudorandomness.

@@

In fact, most existing blind and verifiable protocols for delegation of quantum computation require a notable amount of effort in achieving each property respectively. 
The only successful CVQC protocol~\cite{FOCS:GheVid19} of achieving both so far applies Mahadev's technique~\cite{FOCS:Mahadev18a} to the measurement-based quantum computation, whereas the analysis is still very specific to the construction.
%\Ethan{And it has inverse-poly errors, which makes it not constant-round to achieve negligible errors for $\BQP$, right?}
Could there be a \emph{generic} way to achieve blindness and verifiability for CVQC protocols at the same time, say by leveraging the state-of-the-art CVQC~\cite{FOCS:Mahadev18a} and QFHE~\cite{mahadev_qfhe} protocols?

%\XW{Mention QFHE as a natural attempt, and highlight the problem}

\vspace{2mm} \noindent \textbf{Contribution.} We provide \emph{affirmative} solutions to both of our questions.
In particular, we demonstrate the feasibility of the classical verification of quantum sampling by
constructing a constant-round CVQC protocol for $\SampBQP$, based on the quantum LWE (QLWE) assumption that the learning-with-error problem is hard for BQP machines. 
$\SampBQP$ is the sampling version of $\BQP$ formulated by Aaronson~\cite{aaronson_2013}.
%the class of sampling problem that are approximately solvable by polynomial-time quantum algorithms defined by Aaronson~\cite{aaronson_2013}.
%the sampling version of $\BQP$ formulated by Aaronson~\cite{aaronson_2013}
Formally, $\SampBQP$ consists of sampling problems $(D_x)_{x\in\zo^*}$ that can be approximately sampled by a $\BQP$ machine with an inverse polynomial accuracy. See Section~\ref{sec:samp_definition} for further discussions on $\SampBQP$ and the precise definition of CVQC for $\SampBQP$.   %Namely, $A(x,1^{1/\eps})$ outputs a sample that is $\eps$-close to the distribution $D_x$ in statistical distance.
%, where given an input $x \in \zo^n$, the goal  
%Precisely,
\begin{theorem}[informal] \label{thm:qpip0-informal}
Assuming the QLWE assumption, there exists a four-message CVQC protocol for all sampling problems in $\SampBQP$ with negligible completeness error and computational soundness.
\end{theorem}

@@

%\XW{theorem statement for the second contribution and pointer}

As a simple corollary of combining both results above, we achieve a constant-round blind CVQC protocol for $\SampBQP$. %with negligible completeness error and statistical soundness.  
\begin{theorem}[informal]
        Assuming the QLWE assumption, there exists a blind, four-message CVQC protocol for all sampling problems in $\SampBQP$ with negligible completeness error and computational soundness.
\end{theorem}

@@

%\XW{here for both $\BQP$ and $\SampBQP$}
%\XW{check the para/terminology here; consider adding a theorem statement,or a pointer to the later section}

@@

% \XW{any more to say about parameters, techniques?}
% \XW{anything we want to say about composability?}
% \Ethan{Last time we checked, Vidick might have better composability since he's got some kind of ideal box and simulator-based proof}

% Related work
% \begin{itemize}
%     \item Comparison with related works here?
% \end{itemize}
%       \begin{itemize}
%           \item Sampling, the only paper; how about classical sampling?
%           \item the following for BQP  
%           \item blind and verifiable ~\cite{GV19}; we  constant round; technique-wise very different.
%           \item there is a table in~\cite{Grilo19}. Safe to say ~\cite{GV19} only existing blind protocol in the computational setting?
%           \item all previous either quantum clients, or at least 2 provers.
%           \item the following for blindness
%           \item Mahadev in her thesis~\cite{mahadev_2018} discussed a bit about the relation between verifiability and blindness. She hoped to get verifiability out of blindness by designing some non-malleable QFHE but failed.   
%           \item what's the high-level message we can say here?  Use QFHE in a different way? It is correct that not much work in the classical setting either.  Maybe existing classical work employs the principle but with different implementation.
%           \item old approach, first get blindness  (measurement-based, self-testing), and then try to make it verifiable; our approach, first have a verifiable protocol, and upgrade by a QFHE.
%           \item directly QFHE (blindness) won't give verifiability. some thoughts from Mahadev.
%
%           technical comparison with the past parallel % repetition.
%           \item

\vspace{2mm} \noindent \textbf{Techniques.} Let us revisit Mahadev's CVQC protocol~\cite{FOCS:Mahadev18a} first for some technical background. 
Following~\cite{FOCS:Mahadev18a}, we formally define $\QPIP_{\tau}$ as classes of CVQC protocols where $\tau$ refers to the size of quantum register in the possession of the classical verifier, or equivalently, the limited quantum computation power of the verifier.
It is known that $\BQP$ can be efficiently verified by a classical verifier that can perform a single qubit $X$ or $Z$ measurement~\cite{PhysRevA.93.022326, mf16}, by reducing any $\BQP$ problem to a local Hamiltonian problem where each term consists of  $X$ and $Z$ only. This leads to a $\QPIP_1$ protocol for $\BQP$.
The main contribution of Mahadev~\cite{FOCS:Mahadev18a} can hence be deemed as a way to compile this $\QPIP_1$ protocol into a $\QPIP_0$ protocol (i.e., with a fully classical verifier).

Precisely, to leverage Mahadev's construction, one needs to start with a $\QPIP_1$ protocol with very small completeness and soundness errors, which will then be compiled into a $\QPIP_0$ protocol with a small completeness error, but a close-to-$1/2$ soundness error under the QLWE assumption\footnote{Mahadev only proved soundness error close-to-$3/4$ for her protocol, but it was later proved~\cite{arXiv:ChiaChungYam19, arXiv:AlaChiHun19} that her protocol actually achieves close-to-$1/2$ soundness error}.  
This large soundness error is due to the current structure of Mahadev's protocol that consists of the \emph{testing} round and the \emph{Hadamard} round, each of which happens with a half chance. 
At a high level, the protocol verifies the behavior of the prover in the testing round, while assumes that the prover behaves honestly and all the X-Z measurements are correct in the Hadamard round. 
The soundness is obtained by observing that the dishonest prover cannot cheat in both rounds, while cheating in one round alone is possible which leads to a large soundness error. 
This less desirable soundness error, as well as other parameters, has been subsequently improved in \cite{arXiv:AlaChiHun19, arXiv:ChiaChungYam19} by a parallel repetition of Mahadev's original CVQC protocol in the computational setting.

The first thing when dealing with $\SampBQP$ is to properly define the requirement of CVQC protocols for $\SampBQP$. 
In contrast to the single bit $\mathrm{Accept/Reject}$ information for the decision $\BQP$, any $\SampBQP$ problem needs to always output a sample of the desired distribution.
In the context of CVQC protocols, it means that the protocol should allow the verifier to generate a desired output sample whenever the protocol accepts. 
(Of course, when the prover cheats, the verifier will reject and no further output is required in that case.)
Intuitively, the soundness requires that the verifier $V$ should never be ``cheated'' to accept and output an incorrect sample even when interact with a malicious prover. We formalize this by a strong simulation-based definition (Section~\ref{sec:samp_definition}), where we require that the joint distribution of the decision bit $d \in \set{\Acc, \Rej}$ and the output $z$ (which is $\bot$ when $d = \Rej$) is $\eps$-close (in either statistical or computational sense) to an ``ideal distribution'' $(d,z_{ideal})$, where $z_{ideal}$ is sampled from $D_x$ when $d = \Acc$ and set to $\bot$ when $d = \Rej$.

@@

%We provide a formal definition for CVQC protocols for $\SampBQP$ in Section~\ref{sec:samp_definition} to capture the above intuition. \KM{say the distribution should be correct conditioned on accepts}

%\KM{terminology: testing protocol, test round etc.}

@@

% \XW{revisit Mahadev's approach: (a) QPIP0 to QPIP1; QPIP0 requires very good completeness and soundness. (can say for examples the use of X-Z Hamiltonian, what's the detailed bound transformation?);  the output protocol has poor soundness errors;   (this is still for the decision problem; inefficiency); cite the recent parallel repetition - motivation and result; efficiency- round ? so sequential-repetition is not preferred.}

% \XW{the difference between decision and sampling problems. precision definitions are given in where?}

% \Ethan{Next part needs discussion on context Ie. protocol or local computation}
% Additional challenge comes from modeling and achieving soundness for such a sampling protocol,
% since it is unclear how to apply the usual trick of running many copies of the protocol then applying the Chernoff bound. \XW{too vague to be useful?}

\vspace{2mm} \noindent \textbf{Construction of a $\QPIP_0$ protocol for $\SampBQP$}. 
% \XW{stop here} We will follow the same road map above (i.e., from $\QPIP_1$ to $\QPIP_0$) for $\SampBQP$
% However, since there is no existing $\QPIP_1$ protocol for $\SampBQP$, we make original contributions to both steps as follows:
% \XW{use a different labeling in order to highlight contributions}
Following the aforementioned outline, our construction can be divided into three key technical steps.

% \XW{(1) a basic step for encoding of the history state in the local Hamiltonian problem, and can be tested (so spectral gap). 
% (2) testing the spectral gap and make sure outputing. QPIP0 protocol. (need and how); here also  
% (3) QPIP1 compilation, and also main the efficiency (round?). But this parallel repetition just makes more complicated because of the sampling requirement! 
% }

@@

In the case of $\SampBQP$, one needs to output both Accept/Reject information as well as a sample from the desired distribution if the protocol accepts. 
To that end, one hopes to certify the validity of the entire history state which is the ground state of $H_{\mathrm{circuit}}$, rather than the ground energy of $H_{\mathrm{circuit}}$ only. 
% one still uses $H_{\mathrm{circuit}}$ to certify the validity of the history state.
% However, in this case, one needs to measure on the entire final state of the circuit, rather than a single output qubit,
% which can no longer be encoded solely by the ground energy.
Namely, we want to rule out the existence of any state that is far from the history state but its energy (respect to $H_{\mathrm{circuit}}$) is very close to the ground one. 
One nature approach, also ours, is to make the unique valid history state lie in the ground space of a slightly different local Hamiltonian $H'_{\mathrm{circuit}}$ that has a large \emph{spectral} gap between its ground energy and excited ones.
It is hence guaranteed that any state with a close-to-ground energy must also be close to the history state.
In other words, a certification of the energy $H'_{\mathrm{circuit}}$ could lead to a certification of the history state, which in turn helps us generate the desired sample. 
We construct such $H'_{\mathrm{circuit}}$ from $H_{\mathrm{circuit}}$ by using the \emph{perturbation} technique (e.g.,~\cite{kempe_kitaev_regev_2006}) with further restriction to X/Z terms. (\Cref{sec:LHXZ}.)

%\XW{what's good about Mahadev's protocol; one can measure a lot; but of course, one needs to also test it; for the QPIP1 protocol; to get the QPIP0, the story won't be different.}
% \XW{anything useful here: Our protocol leverages the Hamiltonian model and the computational X-Z measurement from~\cite{FOCS:Mahadev18a}.
% However, a significant amount of new techniques have been developed to deal with the difference between $\SampBQP$ and $\BQP$, which will be highlighted in the technical contribution section. }

\vspace{2mm} \noindent \emph{$\diamond$ A $\QPIP_1$ protocol for $\SampBQP$:} A certification protocol for the history state, however, is insufficient to imply a $\QPIP_1$ protocol for $\SampBQP$ directly. 
Intuitively, we face a similar dilemma as we described above about Mahadev's protocol: we can either employ the certification protocol on $H'_{\mathrm{circuit}}$ to do the test, or to measure the history state to generate the desired outcome, but not both at the same time since the testing protocol would disturb the measurement outcome.
% Specifically, we will certify the energy of $H'_{\mathrm{circuit}}$ to guarantee the underlying state is close to the valid history state.
% However, this procedure could be vastly different from outputting a sample by measuring the final state of $\SampBQP$ circuits.
To resolve that, we design a \emph{cut-and-choose} protocol on multiple copies of the history state so as to separate the testing and the outputting parts on separate copies of history states.

@@

\vspace{2mm} \noindent  \emph{$\diamond$ Compile $\QPIP_1$ into $\QPIP_0$ after parallel repetition}: We now discuss how to use Mahadev's measurement protocol to compile the above $\QPIP_1$ protocol for $\SampBQP$ to a $\QPIP_0$ protocol. As mentioned, a major issue we need to address is that when the verifier $V$ choose to run a testing round, $V$ does not learn a output sample when it accepts.  %(which happens with probability $1/2$),

Specifically, let $\PiNaive$ be an ``intermediate'' $\QPIP_0$ protocol obtained by applying Mahadev's compilation to the above $\QPIP_1$ protocol. Namely, when $V$ chooses to run the Hadamard round, it learns a measurement outcome from the measurement protocol and runs the $\QPIP_1$ verifier to generate a decision and a output sample when accepts. However, when  $V$ chooses to run the testing round, it only decides to accept or reject without being able to output a sample.

@@

Clearly from the construction, the verifier now can output a sample when it decides to accept, and outputs a correct sample when interacting with an honest prover (completeness). The challenge is to show that $\PiSampZ$ is computationally sound. Note that now we are in the computational setting, we cannot use quantum de Finetti theorem as above since it only holds in the information-theoretical setting. Furthermore, parallel repetition for computationally sound protocols are typically difficult to analyze, and known to not always work for protocols with four or more messages~\cite{BIN97,PW12}.

Fortunately, parallel repetition of Mahadev's protocol for $\BQP$ has been analyzed before in ~\cite{arXiv:ChiaChungYam19, arXiv:AlaChiHun19} with interestingly different proofs. Both works showed that $m$-fold parallel repetition reduces the soundness error to $2^{-m}$ by relying on special properties of Mahadev's protocol.

It is worth comparing parallel repetition of Mahadev's protocol for $\BQP$ and $\SampBQP$. For $\BQP$, the verifier simply chooses to run the Hadamard and testing rounds independently for each repetition. In contrast, our $\PiSampZ$ runs the Hadamard round in one repetition and run the testing round in the rest. The reason is that in $\SampBQP$, we do not have a good way to combine two samples learned from two Hadamard-round executions. The difference also reflects in the parameters, where the soundness error decreases exponentially for $\BQP$, but as we see below (and also in the above $\QPIP_1$ protocols), for $\SampBQP$, $m$-fold repetition only decreases the error to $\poly(1/m)$. The difference stems from the fact that there is no generic error reduction method for the sampling problem, whereas one can just conduct a majority vote for decision problems.

We rely on the technique developed in~\cite{arXiv:ChiaChungYam19} to analyze the soundness of $\PiSampZ$. Intuitively, Mahadev's protocol has the property that if a malicious prover $P^*$ knows the answer to pass the testing round, then it must be ``committed'' to a state $\rho$ in the sense that the verifier $V$ learns the $X/Z$ measurement outcome of $\rho$ in the Hadamard round (in a computational sense), and thus, the soundness of the underlying $\QPIP_1$ protocol ensures that $V$ learns a correct sample when it accepts. In short, when $P^*$ knows the testing round answer,  $P^*$ cannot ``cheat'' in the Hadamard round. The work of~\cite{arXiv:ChiaChungYam19} formalizes this intuition by a \emph{partition lemma}, which roughly says that there is an efficient projection that partitions the prover's internal quantum space into a ``known-answer'' subspace $S_1$ that $P^*$ knows the answer to the testing round, and a ``not-known-answer'' subspace $S_0$ that $P^*$ does not know the answer to the testing round. For (efficient) prover's states $\ket{\psi} \in S_1$, the verifier $V$ can produce a sound output in the Hadamard round. For (efficient) prover's states $\ket{\psi} \in S_0$, the verifier $V$ rejects in the testing round with high probability. The work of \cite{arXiv:ChiaChungYam19} used this partition lemma iteratively to each instance of the repetition to show that for a no instance $x$ for any (efficient) prover's state $\ket{\psi}$, the probability that $V$ accepts is at most roughly $2^{-m}$.

>>>>

\section{Introduction}
Can quantum computation, with potential computational advantages that are intractable for classical computers,
be efficiently verified by classical means?
This seeming paradox has been one of the central problems in quantum complexity theory and delegation of quantum computation~\cite{web:Aaronson}.
From a philosophical point of view, this question is also known to have a fascinating connection to the \emph{falsifiability} of quantum mechanics in the high complexity regime~\cite{survey:AV12}.

A complexity theoretic formulation of this problem by Gottesman in 2004~\cite{web:Aaronson} asks about the possibility for an efficient classical verifier (a $\BPP$ machine) to verify the output of an
efficient quantum prover (a $\BQP$ machine).
In the absence of techniques for directly tackling this question, earlier feasibility results on this problem have been focusing on two weaker formulations.
The first type of feasibility results (e.g.,~\cite{BFK09,arXiv:ABOEM17,FK17,mf16}) considers the case where the $\BPP$ verifier is equipped with limited quantum power.
The second type of feasibility results (e.g,~\cite{Nat:RUV13, CGJV19, Gheorghiu_2015, HPF15})
considers a $\BPP$ verifier interacting with at least two entangled, non-communicating quantum provers.
In a recent breakthrough, Mahadev~\cite{FOCS:Mahadev18a} proposed the first protocol of Classical Verification of Quantum Computation (CVQC) whose soundness is based on a widely recognized computational assumption that the learning with error (LWE) problem~\cite{JACM:Regev09} is hard for $\BQP$ machines.
The technique invented therein has inspired many  subsequent developments of CVQC protocols with improved parameters and functionality (e.g.,~\cite{FOCS:GheVid19,arXiv:AlaChiHun19,arXiv:ChiaChungYam19}).
On the other side, there are known complexity-theoretic limitations on the feasibility of blind CVQC in the information-theoretical setting (e.g.~\cite{aaronson_et_al:LIPIcs:2019:10582}).

@@

It is worthwhile noting that there is a simple reduction of the delegation of \emph{classical} sampling problems to decision ones: for example, the verifier can fix a seed for pseudo-randomness, send it to the prover, and then the sampling outcome can be computed  bit-by-bit under this pseudo-random seed in a deterministic way. As such, classical literature of delegation of computation primarily focuses on delegation of decision problems.

Unfortunately, it is unclear about how to make this de-randomization trick work for the delegation of $\SampBQP$, 
as randomness in quantum computation inherently comes from quantum mechanics rather than classical probability theory. 
Thus, it seems that the delegation of $\SampBQP$ needs to take a different technical route.
One natural starting point is Mahadev's CVQC protocol~\cite{FOCS:Mahadev18a} for $\BQP$. 
Indeed, some nice features of Mahadev's protocol (e.g., allowing X-Z measurements on any qubit) suggest the feasibility of generating  multi-bit measurement outcomes with the protocol. 
However, as highlighted below in the technical introduction section, there are also important drawbacks of Mahadev's protocol in its current form, which makes it hard to allow a CVQC protocol for $\SampBQP$ with desired performance. As we shall see, such difficulties do not hold for the decision problem, which demonstrates an important difference between CVQC protocols for the decision and the sampling problems.

@@

In fact, most existing blind and/or verifiable protocols for delegation of quantum computation require a notable amount of effort in achieving each property respectively. 
The only successful CVQC protocol~\cite{FOCS:GheVid19} of achieving both so far applies Mahadev's technique~\cite{FOCS:Mahadev18a} to the measurement-based quantum computation, whereas the analysis is still very specific to the construction.
Could there be a \emph{generic} way to achieve blindness and verifiability for CVQC protocols at the same time, say by leveraging the state-of-the-art CVQC~\cite{FOCS:Mahadev18a} and QFHE~\cite{mahadev_qfhe} protocols?

\vspace{2mm} \noindent \textbf{Contribution.} We provide \emph{affirmative} solutions to both of our questions.
In particular, we demonstrate the feasibility of the classical verification of quantum sampling by
constructing a constant-round CVQC protocol for $\SampBQP$, based on the quantum LWE (QLWE) assumption that the learning-with-error problem is hard for BQP machines. 
$\SampBQP$ is the sampling version of $\BQP$ formulated by Aaronson~\cite{aaronson_2013}.
Formally, $\SampBQP$ consists of sampling problems $(D_x)_{x\in\zo^*}$ that can be approximately sampled by a $\BQP$ machine with an inverse polynomial accuracy. See Section~\ref{sec:samp_definition} for further discussions on $\SampBQP$ and the precise definition of CVQC for $\SampBQP$.   \begin{theorem}[informal] \label{thm:qpip0-informal}
Assuming the QLWE assumption, there exists a four-message CVQC protocol for all sampling problems in $\SampBQP$ with negligible completeness error and computational soundness.
\end{theorem}

@@

As a simple corollary of combining both results above, we achieve a constant-round blind CVQC protocol for $\SampBQP$. 
\begin{theorem}[informal]
        Assuming the QLWE assumption, there exists a blind, four-message CVQC protocol for all sampling problems in $\SampBQP$ with negligible completeness error and computational soundness.
\end{theorem}

@@

@@

\vspace{2mm} \noindent \textbf{Techniques.} Let us revisit Mahadev's CVQC protocol~\cite{FOCS:Mahadev18a} first for some technical background. 
Following~\cite{FOCS:Mahadev18a}, we formally define $\QPIP_{\tau}$ as classes of CVQC protocols where $\tau$ refers to the size of (local) quantum memory in the possession of the classical verifier, or equivalently, the limited quantum power of the verifier\footnote{Intuitively, the size of local quantum memory limits the size of qubits that a general (entangled) quantum operation can be operated on at the same time. However, the verifier is still able to operate on all the qubits in a streaming fashion: i.e., receive the qubits sent from the pover and discard some local qubits in the case of overflow.}. 
A precise definition is given in Section~\ref{sec:qpip_def}. 
It is known that $\BQP$ can be efficiently verified by a classical verifier that can perform a single qubit $X$ or $Z$ measurement~\cite{PhysRevA.93.022326, mf16}, by reducing any $\BQP$ problem to a local Hamiltonian problem where each term consists of  $X$ and $Z$ only. This leads to a $\QPIP_1$ protocol for $\BQP$.
The main contribution of Mahadev~\cite{FOCS:Mahadev18a} can hence be deemed as a way to compile this $\QPIP_1$ protocol into a $\QPIP_0$ protocol (i.e., with a fully classical verifier).

Precisely, to leverage Mahadev's construction, one needs to start with a $\QPIP_1$ protocol with very small completeness and soundness errors, which will then be compiled into a $\QPIP_0$ protocol with a small completeness error, but a close-to-$1/2$ soundness error under the QLWE assumption\footnote{Mahadev only proved soundness error close-to-$3/4$ for her protocol, but it was later proved~\cite{arXiv:ChiaChungYam19, arXiv:AlaChiHun19} that her protocol actually achieves close-to-$1/2$ soundness error.}.  
This large soundness error is due to the current structure of Mahadev's protocol that consists of the \emph{testing} round and the \emph{Hadamard} round, each of which happens with a half chance. 
At a high level, the protocol verifies the behavior of the prover in the testing round, while assumes that the prover behaves honestly and all the X-Z measurements are correct in the Hadamard round. 
The soundness is obtained by observing that the dishonest prover cannot cheat in both rounds, while cheating in one round alone is possible which leads to a large soundness error. 
This less desirable soundness error, as well as other parameters, has been subsequently improved in \cite{arXiv:AlaChiHun19, arXiv:ChiaChungYam19} by a parallel repetition of Mahadev's original CVQC protocol in the computational setting.

The first and important step in the case of $\SampBQP$ is to identify an appropriate definition of corresponding CVQC protocols. In contrast to the single bit $\mathrm{Accept/Reject}$ information for the decision $\BQP$, any $\SampBQP$ problem needs to always output a sample of the desired distribution.
In the context of CVQC protocols, it means that the protocol should allow the verifier to generate a desired output sample whenever the protocol accepts. 
(Of course, when the prover cheats, the verifier will reject and no further output is required in that case.)
Intuitively, the soundness requires that the verifier $V$ should never be ``cheated'' to accept and output an incorrect sample even when interact with a malicious prover. We formalize this by a strong \emph{simulation-based} definition (Section~\ref{sec:samp_definition}), where we require that the joint distribution of the decision bit $d \in \set{\Acc, \Rej}$ and the output $z$ (which is $\bot$ when $d = \Rej$) is $\eps$-close (in the either statistical or computational sense) to an ``ideal distribution'' $(d,z_{ideal})$, where $z_{ideal}$ is sampled from $D_x$ when $d = \Acc$ and set to $\bot$ when $d = \Rej$.

@@

@@

\vspace{2mm} \noindent \textbf{Construction of a $\QPIP_0$ protocol for $\SampBQP$}. 
Following the aforementioned outline, our construction can be divided into three key technical steps.

@@

In the case of $\SampBQP$, one needs to output both Accept/Reject information as well as a sample from the desired distribution if the protocol accepts. 
To that end, one hopes to certify the validity of the entire history state which is the ground state of $H_{\mathrm{circuit}}$, rather than the ground energy of $H_{\mathrm{circuit}}$ only. 
% one still uses $H_{\mathrm{circuit}}$ to certify the validity of the history state.
% However, in this case, one needs to measure on the entire final state of the circuit, rather than a single output qubit,
% which can no longer be encoded solely by the ground energy.
Namely, we want to rule out the existence of any state that is far from the history state but its energy (respect to $H_{\mathrm{circuit}}$) is very close to the ground one. 
One nature approach, also ours, is to make the unique valid history state lie in the ground space of a slightly different local Hamiltonian $H'_{\mathrm{circuit}}$ that has a large \emph{spectral} gap between its ground energy and excited ones.
It is hence guaranteed that any state with a close-to-ground energy must also be close to the history state.
In other words, a certification of the energy $H'_{\mathrm{circuit}}$ could lead to a certification of the history state, which in turn helps us generate the desired sample. 
We construct such $H'_{\mathrm{circuit}}$ from $H_{\mathrm{circuit}}$ by using the \emph{perturbation} technique (e.g.,~\cite{kempe_kitaev_regev_2006}) with a further restriction to X/Z terms. (\Cref{sec:LHXZ}.)

\vspace{2mm} \noindent \emph{$\diamond$ A $\QPIP_1$ protocol for $\SampBQP$:} A certification protocol for the history state, however, is insufficient to imply a $\QPIP_1$ protocol for $\SampBQP$ directly. 
Intuitively, we face a similar dilemma as we described above about Mahadev's protocol: we can either employ the certification protocol on $H'_{\mathrm{circuit}}$ to do the test, or to measure the history state to generate the desired outcome, but not both at the same time since the testing protocol would disturb the measurement outcome.
To resolve that, we design a \emph{cut-and-choose} protocol on multiple copies of the history state so as to separate the testing and the outputting parts on separate copies of history states.

@@

\vspace{2mm} \noindent  \emph{$\diamond$ Compile $\QPIP_1$ into $\QPIP_0$ after parallel repetition}: We now discuss how to use Mahadev's measurement protocol to compile the above $\QPIP_1$ protocol for $\SampBQP$ to a $\QPIP_0$ protocol. As mentioned, a major issue we need to address in Mahadev's original construction is that when the verifier $V$ chooses to run a testing round, $V$ does not learn an output sample when it accepts.  %(which happens with probability $1/2$),

Specifically, let $\PiNaive$ be an ``intermediate'' $\QPIP_0$ protocol obtained by applying Mahadev's compilation to the above $\QPIP_1$ protocol. Namely, when $V$ chooses to run the Hadamard round, it could learn a measurement outcome from the measurement protocol and be able to run the $\QPIP_1$ verifier to generate a decision and an output sample when accepts. However, when  $V$ chooses to run the testing round, it only decides to accept/reject without being able to output a sample.

@@

Clearly from the construction, the verifier now can output a sample when it decides to accept, and output a correct sample when interacting with an honest prover (completeness). The challenge is to show that $\PiSampZ$ is computationally sound. Note that now we are in the computational setting, we cannot use quantum de Finetti theorem as above since it only holds in the information-theoretical setting. Furthermore, parallel repetition for computationally sound protocols are typically difficult to analyze, and known to not always work for protocols with four or more messages~\cite{BIN97,PW12}.

Fortunately, parallel repetition of Mahadev's protocol for $\BQP$ has been analyzed before in ~\cite{arXiv:ChiaChungYam19, arXiv:AlaChiHun19} with interestingly different proofs. Both works showed that $m$-fold parallel repetition reduces the soundness error to $2^{-m}$ for $\BQP$ by relying on special properties of Mahadev's protocol.

It is worth comparing parallel repetition of Mahadev's protocol for $\BQP$ and $\SampBQP$. For $\BQP$, the verifier simply chooses to run the Hadamard and testing rounds independently for each repetition. In contrast, our $\PiSampZ$ runs the Hadamard round in one repetition and run the testing round in the rest. The reason is that in $\SampBQP$, as well as generically in sampling problems, there is no known approach to combine two samples learned from two Hadamard-round executions to generate one sample with much better parameters, i.e., no generic error reduction method for the sampling problem. 
In contrast, the error reduction for decision problems can be done with the majority vote. 
As a result, while the soundness error decreases exponentially for $\BQP$, as we see below (and also in the above $\QPIP_1$ protocols), for $\SampBQP$, $m$-fold repetition only decreases the error to $\poly(1/m)$.

We rely on the technique developed in~\cite{arXiv:ChiaChungYam19} to analyze the soundness of $\PiSampZ$. Intuitively, Mahadev's protocol has the property that if a malicious prover $P^*$ knows the answer to pass the testing round, then it must be ``committed'' to a state $\rho$ in the sense that the verifier $V$ learns the $X/Z$ measurement outcome of $\rho$ in the Hadamard round (in a computational sense). Thus, the soundness of the underlying $\QPIP_1$ protocol ensures that $V$ learns a correct sample when it accepts. In short, when $P^*$ knows the testing round answer,  $P^*$ cannot ``cheat'' in the Hadamard round. The work of~\cite{arXiv:ChiaChungYam19} formalizes this intuition by a \emph{partition lemma}, which roughly says that there is an efficient projection that partitions the prover's internal quantum space into a ``known-answer'' subspace $S_1$ where $P^*$ knows the answer to the testing round, and a ``not-known-answer'' subspace $S_0$ where $P^*$ does not know the answer to the testing round. For (efficient) prover's states $\ket{\psi} \in S_1$, the verifier $V$ can produce a sound output in the Hadamard round. For (efficient) prover's states $\ket{\psi} \in S_0$, the verifier $V$ rejects in the testing round with high probability. The work of \cite{arXiv:ChiaChungYam19} used this partition lemma iteratively to each instance of the repetition to show that for a no instance $x$ for any (efficient) prover's state $\ket{\psi}$, the probability that $V$ accepts is at most roughly $2^{-m}$.
